Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-10 paper/paper.tex
almost final plots

\subsection{Results}
\newcommand{\hgt}{3cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=gc_bench.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=gc_bench_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\end{tabular}\vskip -10mm
 \caption{Memory usage  snapshot of  programs.  The  blue and  the red
   curves indicate the  number of cons cells in  the active semi-space
   for  LGC and  RGC  respectively.  The  black  curve represents  the
   number of reachable cells and  the grey curve represents the number
   of cells that are actually live  (of which liveness analysis does a
   static approximation).  x-axis is  the time  measured in  number of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

@@

\section{Related Work} 
Although augmenting  garbage collection with liveness  information has
been  studied earlier,  it was  mainly  in the  context of  imperative
languages~\cite{Albert,  Hirzel}.  Due  to   the  presence  of  global
variables  and mutation,  practical utility  of simple  liveness based
techniques are found to be inefficient.

In the space  of functional languages, improving  space efficiency has
been  mainly  studied  as  a compile  time  activity.  Either  through
rewriting                methods                such                as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing  analysis   based  reallocation~\cite{jones89compile},  region
based  analysis~\cite{tofte98region} or  through insertion  of compile
time              nullifying              statements              such
as~\cite{inoue88analysis,lee05static,Hamilton}.    The  compile   time
marking approaches all rely on an efficient and precise alias analysis
and in the absence of it cannot provide notable improvement.

The closest  to our approach  is the liveness based  garbage collector
implemented in~\cite{asati14lgc}. We extend  their work to handle lazy
evaluation and closures.

%% A similar approach has been suggested in~\cite{ONeill} where they suggest augmenting garbage collection with extra information using {\it simplifiers}.

%% Previous  attempts  to increase  the  space  efficiency of  functional
%% programs  by  additional  reclamation  of memory  fall  in  two  broad
%% categories. In the first, the program itself is instrumented to manage
%% reclamation and reallocation without the aid of the garbage collector.
%% Such      attempts      include:      sharing      analysis      based
%% reallocation~\cite{jones89compile},                      deforestation
%% techniques~\cite{wadler88deforest,gill93ashort,chitil99deforest},
%% methods  based  on   linear  logic~\cite{hofmann00linear}  and  region
%% analysis~\cite{tofte98region}.   Closer  to  our approach,  there  are
%% methods   that  enable   the   garbage  collector   to  collect   more
%% garbage~\cite{inoue88analysis,lee05static}  by  explicitly  nullifying
%% pointers  that are  not  live.  However,  the  nullification, done  at
%% compile  time,  requires sharing  (alias)  analysis.   Our method,  in
%% contrast, does not require sharing  because of the availability of the
%% heap itself  at runtime.  To  the best of  our knowledge, this  is the
%% first attempt  at liveness-based  marking of  the heap  during garbage
%% collection.

>>>>

\subsection{Results}
\newcommand{\hgt}{3cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory  usage  for (a)  full  benchmarks  and (b)  for  some
   interesting part  of the benchmarks.   The blue and the  red curves
   indicate the number of cons cells  in the active semi-space for RGC
   and LGC  respectively.  The  black curve  represents the  number of
   reachable cells and  the grey curve represents the  number of cells
   that are  actually live (of  which liveness analysis does  a static
   approximation).   x-axis   is  the  time  measured   in  number  of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

@@

\section{Related Work} 
\cred{Need to  rewrite this  entire section  need to  probably mention
  some of  these papers at  least i  found by googling  liveness based
  GC,\\ http://www.cs.hmc.edu/~oneill/papers/Simplifiers-MSPC.pdf\\ http://costa.ls.fi.upm.es/papers/costa/AlbertGGZ12.pdf\\ http://hirzels.com/martin/papers/ecoop01-gc-liveness.pdf\\ }
Previous  attempts  to increase  the  space  efficiency of  functional
programs  by  additional  reclamation  of memory  fall  in  two  broad
categories. In the first, the program itself is instrumented to manage
reclamation and reallocation without the aid of the garbage collector.
Such      attempts      include:      sharing      analysis      based
reallocation~\cite{jones89compile},                      deforestation
techniques~\cite{wadler88deforest,gill93ashort,chitil99deforest},
methods  based  on   linear  logic~\cite{hofmann00linear}  and  region
analysis~\cite{tofte98region}.   Closer  to  our approach,  there  are
methods   that  enable   the   garbage  collector   to  collect   more
garbage~\cite{inoue88analysis,lee05static}  by  explicitly  nullifying
pointers  that are  not  live.  However,  the  nullification, done  at
compile  time,  requires sharing  (alias)  analysis.   Our method,  in
contrast, does not require sharing  because of the availability of the
heap itself  at runtime.  To  the best of  our knowledge, this  is the
first attempt  at liveness-based  marking of  the heap  during garbage
collection.
